---
title: "Non-Life Insurance - Final project"
output: html_document
author: "Klemen Hovnik and Manca Strgar"
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages,  include=FALSE, warning = FALSE, message = FALSE}
#package instalation:
packages <- c("tidyverse", "mgcv", "evtree", "classInt", "rgdal", "RColorBrewer", "grid", "gridExtra", "visreg", "sf", "tmap", "rgeos", "mapview", "leaflet", "rmarkdown", "ggplot2", "kableExtra")
suppressMessages(packages <- lapply(packages, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x)
    library(x, character.only = TRUE)
  }
  else {
  message("Everything is set up correctly. You are ready to go.")
}
}))

#libraries:
library(dplyr)
library(knitr)
library(rmarkdown)
library(tidyverse)
library(ggplot2)
library(graphics)
library(here)
library(gridExtra)
library(mgcv)
library(evtree)
library(classInt)
library(rgdal)
library(RColorBrewer)
library(grid)
library(visreg)
library(sf)
library(tmap)
library(rgeos)
library(mapview)
library(leaflet)
library(kableExtra)


```
PART 1
=======

# 1 Importing dataset

First, we need to import data from the file `Assignment.csv` into R. The file contains 163.657 rows with 16 variables. The column `nbrtotc` shows total number of claims during the period of exposure. 
Let's have a look at our dataset:

```{r, echo=F}
mtpl_orig <- read.csv("Assignment.csv", header = TRUE)
mtpl_orig=as_tibble(mtpl_orig)
#kable(head(mtpl_orig))
#str(mtpl_orig)
#summary(mtpl_orig)

```

We could rename colums in our dataset to make it easier to work with. We will rename the column `nbrtotc` into `nclaims` as the number of claims, and `duree` into `expo` as exposure. For easier programming we will also rename `ageph` and `sexp` into `age` and `sex`. We aslo renamed `chargtot` into `amount`. We will also delete colums which are not important for our analysis rightnow. 
```{r, echo=F}
mtpl <- mtpl_orig %>%
  # rename all columns 
  rename_all(function(.name) {
    .name %>% 
      # replace all names with the lowercase versions
      tolower 
    })
mtpl <- rename(mtpl, age= ageph, sex = sexp, nclaims = nbrtotc, expo = duree,
               amount = chargtot)
mtpl$amount[mtpl$amount == 0] <- NA 
#kable(head(mtpl))
```


# 3.2 The construction of a (technical) tariff structure (Manca)
Firs lets build a Poisson GLM for numberof claims, so we are building frequency model. For frequency fitting is the best to use Poisson distribution or Negative Binomial distribution. We will built a GLM for number of claims as a function of a covariance gender, we will also include the log of exposure as an offset.
```{r, echo=F}
freq_glm_sex <- glm(nclaims ~ sex, offset =log(expo), family =poisson(link = "log"),
                  data = mtpl)

freq_glm_sex %>% broom::tidy()
freq_glm_sex %>% broom::augment(type.predict = "response")
emp_freq_female <- exp(coef(freq_glm_sex)[1])
emp_freq_male <- exp(coef(freq_glm_sex)[1]+coef(freq_glm_sex)[2])    #TU SEM POPRAVO GLM_1)[1] V GLM_SEX)[1]

c(emp_freq_female, emp_freq_male)
```
If we now look back at the begining of our report, we analysed empirical frequency grouped by the gender. We can see, that we got the same result for empirical frequency by gender, as we did in the first part of the report. this is because we constructed the GLM with only one covariance (sex).
In RStudio we also have a `predict` function that alows us to use our GLM model on new data frames. We can now dafine two new data frames, one being male drivers with exposure 1 and the other being the female drivers with exposure 1. And we can use `predict` function on these two new data frames
```{r, echo=F}
male_driver <- data.frame(expo = 1, sex = "Male")
female_driver <- data.frame(expo = 1, sex = "Female")
c(predict(freq_glm_sex, newdata = male_driver,    #TU SEM POPRAVO _1 V _SEX
       type = "response"),
predict(freq_glm_sex, newdata = female_driver, 
       type = "response"))
```
Now we will build a model for sevirity information. For severity, the best distributions for fitting are Gamma and Log-normal distribution. We will firstly do Gamma GLM. So for sevirity we will analyse variable `amount` in our mtpl data. We will again use only one explenatory variable sex. 
```{r, echo=F}
mtpl$amount[mtpl$amount == 0] <- NA     
sev_glm_1 <- glm(amount ~ sex, offset =log(expo),family = Gamma(link = "log"), data =mtpl)
sev_glm_1
sev_glm_1 %>% broom::tidy()
sev_glm_1 %>% broom::augment(type.predict = "response")
```
Now lets try some other model than GLM. We can model frequency and sevirity also with Generalized Additive Models (GAM).These models are exactly like the formula for a GLM except that smooth term, `s`,  can be added to the right hand side to specify that the linear predictor depends on smooth functions of predictors (or linear functionals of these).
```{r, echo=F}
freq_gam_age <- gam(nclaims ~ s(age, bs ="cr"),   #s = cr, denotes cubic regression splines
                  method = "REML",
                  family = poisson(link = "log"), 
                  data =mtpl)
freq_gam_sex <- gam(nclaims ~ sex,
                  method = "REML",
                  family = poisson(link = "log"), 
                  data =mtpl)




freq_gam_sex %>% broom::augment(type.predict = "response")
emp_freq_female_gam <- exp(coef(freq_gam_sex)[1])
emp_freq_male_gam <- exp(coef(freq_gam_sex)[1]+coef(freq_gam_sex)[2])   

c(emp_freq_female_gam, emp_freq_male_gam)



mtpl$amount[mtpl$amount == 0] <- NA     
sev_gam_1 <- gam(amount ~ sex, method = "REML", family = Gamma(link = "log"), data =mtpl)
#AIC(sev_glm_1) #309561
#AIC(sev_gam_1) #303662

sev_gam_1 %>% broom::augment(type.predict = "response")
```
Now let's make the same 20 models as before but use GAM regression instead of GLM. 
```{r, echo=F, include = False}
freq_gam_1 <- gam(nclaims ~ s(age) + sex + s(codposs) + agecar + fuelc + split + usec + fleetc + sportc + coverp + powerc, method = "REML", family =poisson(link = "log"),data = mtpl)
AIC_freq_gam_1 <- AIC(freq_gam_1) #AIC = 126317

freq_gam_2 <- gam(nclaims ~ s(age) + s(codposs) + agecar + fuelc + split + usec + fleetc + sportc + coverp + powerc, method = "REML", family =poisson(link = "log"),data = mtpl)
AIC_freq_gam_2 <-AIC(freq_gam_2) #AIC = 126318.87  model 1 is a better fit

freq_gam_3 <- gam(nclaims ~ s(age) + sex + s(codposs) + agecar + fuelc + split +  fleetc + sportc + coverp + powerc, method = "REML", family =poisson(link = "log"),data = mtpl)
AIC_freq_gam_3 <- AIC(freq_gam_3) #AIC = 126316.1 this model is a better fit

freq_gam_4 <- gam(nclaims ~ s(age) + sex + age:sex + s(codposs) + agecar + fuelc + split + fleetc + sportc + coverp + powerc,  method = "REML", family =poisson(link = "log"), data = mtpl)
AIC_freq_gam_4 <- AIC(freq_gam_4) #AIC = 126284.6 this model is a better fit

freq_gam_5 <- gam(nclaims ~ s(age) + sex + age:sex + age:powerc + s(codposs) + agecar + fuelc + split +  fleetc + sportc + coverp + powerc, method = "REML", family =poisson(link = "log"),data = mtpl)
AIC_freq_gam_5 <- AIC(freq_gam_5) #AIC = 126286 model 4 is  a better fit

freq_gam_6 <- gam(nclaims ~ s(age) + sex + age:sex +  s(codposs) + agecar + fuelc + split + sportc + coverp + powerc, method = "REML", family =poisson(link = "log"),data = mtpl)
AIC_freq_gam_6 <- AIC(freq_gam_6) #AIC = 126284.8 model 4 is  a better fit

freq_gam_7 <- gam(nclaims ~ s(age) + sex + age:sex +  s(codposs) + agecar + fuelc + split + fleetc + sportc +  powerc, method = "REML", family =poisson(link = "log"),data = mtpl)
AIC_freq_gam_7 <- AIC(freq_gam_7) #AIC = 126313.4 model 4 is  a better fit

freq_gam_8 <- gam(nclaims ~ s(age) + sex + age:sex + s(codposs) + agecar + split + fleetc + sportc + coverp + powerc, method = "REML", family =poisson(link = "log"),data = mtpl)
AIC_freq_gam_8 <- AIC(freq_gam_8) #AIC = 126428.6 model 4 is  a better fit

freq_gam_9 <- gam(nclaims ~ s(age) + sex + age:sex + s(codposs) + fuelc + split + fleetc + sportc + coverp + powerc, method = "REML", family =poisson(link = "log"),data = mtpl)
AIC_freq_gam_9 <- AIC(freq_gam_9) #AIC = 126286 model 4 is  a better fit

##freq_gam_10 <- gam(nclaims ~ s(age) + sex + age:sex + + agecar + fuelc + split + fleetc + sportc + coverp + powerc, method = ##"REML", family =poisson(link = "log"),data = mtpl)
##AIC(freq_gam_10) #AIC = 126314.5 model 4 is  a better fit
AIC_freq_gam_10 <- 126314.5

freq_gam_11 <- gam(nclaims ~ s(age) + sex + age:sex + s(codposs) + agecar + fuelc + split + fleetc + coverp + powerc, method = "REML", family =poisson(link = "log"),data = mtpl)
AIC_freq_gam_11 <- AIC(freq_gam_11) #AIC = 126282.6 This model is a better fit

freq_gam_12 <- gam(nclaims ~ s(age) + sex + age:sex + s(codposs) + agecar + fuelc + split + fleetc + coverp, method = "REML", family =poisson(link = "log"),data = mtpl)
AIC_freq_gam_12 <- AIC(freq_gam_12) #AIC = 126314 model 11 is a better fit

#freq_gam_13 <- gam(nclaims ~ s(age) + sex + age:sex, method = "REML", family =poisson(link = "log"),data = mtpl)
#AIC(freq_gam_13) #AIC =  12650573.1 model 11 is a better fit   needs too much time
AIC_freq_gam_13 <- 126310

freq_gam_14 <- gam(nclaims ~ s(age) + sex + age:sex + s(codposs) + agecar + fuelc + fleetc + coverp + powerc, method = "REML", family =poisson(link = "log"),data = mtpl)
AIC_freq_gam_14 <- AIC(freq_gam_14) #AIC = 126505 model 11 is a better fit

freq_gam_15 <- gam(nclaims ~ s(age) + sex + age:sex + s(codposs) + agecar + fuelc + agecar:fuelc + split + fleetc + coverp + powerc, method = "REML", family =poisson(link = "log"),data = mtpl)
AIC_freq_gam_15 <- AIC(freq_gam_15) #AIC = 126267.3 this model is a better fit

freq_gam_16 <- gam(nclaims ~ s(age) + sex + age:sex + s(codposs) + agecar + fuelc + agecar:fuelc + split + fleetc + coverp + powerc + sex:powerc, method = "REML", family =poisson(link = "log"),data = mtpl)
AIC_freq_gam_16 <- AIC(freq_gam_16) #AIC = 126266.7 this model is a better fit

#freq_gam_17 <- gam(nclaims ~ s(age) + sex + s(codposs) + agecar + fuelc + split + fleetc + coverp + powerc +  agecar:fuelc + #age:sex + sex:powerc + age:codposs, method = "REML", family =poisson(link = "log"),data = mtpl)
#AIC(freq_gam_17) #  model 16 is a better fit
AIC_freq_gam_17 <- 126270

freq_gam_18 <- gam(nclaims ~ s(age) + sex + s(codposs) + agecar + fuelc + split + fleetc + coverp + powerc +  agecar:fuelc + age:sex + sex:powerc + age:split, method = "REML", family =poisson(link = "log"),data = mtpl)
AIC_freq_gam_18 <- AIC(freq_gam_18) #AIC = 126257 this model is a better fit

freq_gam_19 <- gam(nclaims ~ s(age) + sex + s(codposs) + agecar + fuelc + split + fleetc + coverp + powerc +  agecar:fuelc + age:sex + sex:powerc + age:split + age:agecar, method = "REML", family =poisson(link = "log"),data = mtpl)
AIC_freq_gam_19 <- AIC(freq_gam_19) #AIC = 126251.9 this model is a better fit

freq_gam_20 <- gam(nclaims ~ s(age) + sex + s(codposs) + agecar + fuelc + split + fleetc + coverp + powerc +  agecar:fuelc + age:sex + sex:powerc + age:split + age:agecar + fuelc:coverp, method = "REML", family =poisson(link = "log"),data = mtpl)
AIC_freq_gam_20 <- AIC(freq_gam_20) #AIC = 126254 model 19 is a better fit

AIC(freq_gam_20) #AIC = 126106.2 model 19 is a better fit
x <- c(1,2,3,4,5,6,7,8,9,11,12,14,15,16,18,19,20)
AIC_freq_gam <- data.frame(model_name = numeric(0),  AIC = numeric(0))
for(i in x ){
  AIC_freq_gam[i,] <- c(paste("freq_gam_",i, sep =""),round(AIC(get(paste("freq_gam_",i, sep =""))),1) ) 
}

```
Fitting severity model
```{r, echo=F}

sev_gam_1 <- gam(amount ~ s(age) + sex + s(codposs) + agecar + fuelc + split + usec + fleetc + sportc + coverp + powerc, family =Gamma(link = "log"),
                  data = mtpl)
AIC(sev_gam_1) #AIC = 302864.4

sev_gam_2 <- gam(amount ~ s(age) + s(codposs) + agecar + fuelc + split + usec + fleetc + sportc + coverp + powerc, method = "REML", family =Gamma(link = "log"),data = mtpl)
AIC_sev_gam_2 <- AIC(sev_gam_2) #AIC = 302890.4  model 1 is a better fit

sev_gam_3 <- gam(amount ~ s(age) + sex + s(codposs) + agecar + fuelc + split +  fleetc + sportc + coverp + powerc, method = "REML", family = Gamma(link = "log"),data = mtpl)
AIC_sev_gam_3 <- AIC(sev_gam_3) #AIC = 302865 this model is a better fit

sev_gam_4 <- gam(amount ~ s(age) + sex + age:sex + s(codposs) + agecar + fuelc + split + fleetc + sportc + coverp + powerc,  method = "REML", family =Gamma(link = "log"), data = mtpl)
AIC_sev_gam_4 <- AIC(sev_gam_4) #AIC = 302864.7 this model is a better fit

sev_gam_5 <- gam(amount ~ s(age) + sex + age:sex + age:powerc + s(codposs) + agecar + fuelc + split +  fleetc + sportc + coverp + powerc, method = "REML", family =Gamma(link = "log"),data = mtpl)
AIC_sev_gam_5 <- AIC(sev_gam_5) #AIC = 302857.9 model 5 is  a better fit

sev_gam_6 <- gam(amount ~ s(age) + sex + age:sex +  s(codposs) + agecar + fuelc + split + sportc + coverp + powerc, method = "REML", family =Gamma(link = "log"),data = mtpl)
AIC_sev_gam_6 <- AIC(sev_gam_6) #AIC = 302868.9 model 5 is  a better fit

sev_gam_7 <- gam(amount ~ s(age) + sex + age:sex +  s(codposs) + agecar + fuelc + split + fleetc + sportc +  powerc, method = "REML", family =Gamma(link = "log"),data = mtpl)
AIC_sev_gam_7 <- AIC(sev_gam_7) #AIC = 302952 model 5 is  a better fit

sev_gam_8 <- gam(amount ~ s(age) + sex + age:sex + s(codposs) + agecar + split + fleetc + sportc + coverp + powerc, method = "REML", family =Gamma(link = "log"),data = mtpl)
AIC_sev_gam_8 <- AIC(sev_gam_8) #AIC = 302897 model 5 is  a better fit

sev_gam_9 <- gam(amount ~ s(age) + sex + age:sex + s(codposs) + fuelc + split + fleetc + sportc + coverp + powerc, method = "REML", family =Gamma(link = "log"),data = mtpl)
AIC_sev_gam_9 <- AIC(sev_gam_9) #AIC = 302937 model 5 is  a better fit

sev_gam_10 <- gam(amount ~ s(age) + sex + age:sex + + agecar + fuelc + split + fleetc + sportc + coverp + powerc, method = "REML", family =Gamma(link = "log"),data = mtpl)
AIC(sev_gam_10) #AIC = 302981 model 5 is  a better fit

sev_gam_11 <- gam(amount ~ s(age) + sex + age:sex + s(codposs) + agecar + fuelc + split + fleetc + coverp + powerc, method = "REML", family =Gamma(link = "log"),data = mtpl)
AIC_sev_gam_11 <- AIC(sev_gam_11) #AIC = 302865.5  model 5 is a better fit

sev_gam_12 <- gam(amount ~ s(age) + sex + age:sex + s(codposs) + agecar + fuelc + split + fleetc + coverp, method = "REML", family =Gamma(link = "log"),data = mtpl)
AIC_sev_gam_12 <- AIC(sev_gam_12) #AIC = 302887 model 5 is a better fit

sev_gam_13 <- gam(amount ~ s(age) + sex + age:sex, method = "REML", family =Gamma(link = "log"),data = mtpl)
AIC(sev_gam_13) #AIC =  303408 model 5 is a better fit   needs too much time

sev_gam_14 <- gam(amount ~ s(age) + sex + age:sex + s(codposs) + agecar + fuelc + fleetc + coverp + powerc, method = "REML", family =Gamma(link = "log"),data = mtpl)
AIC_sev_gam_14 <- AIC(sev_gam_14) 

sev_gam_15 <- gam(amount ~ s(age) + sex + age:sex + s(codposs) + agecar + fuelc + agecar:fuelc + split + fleetc + coverp + powerc, method = "REML", family =Gamma(link = "log"),data = mtpl)
AIC_sev_gam_15 <- AIC(sev_gam_15) 

sev_gam_16 <- gam(amount ~ s(age) + sex + age:sex + s(codposs) + agecar + fuelc + agecar:fuelc + split + fleetc + coverp + powerc + sex:powerc, method = "REML", family =Gamma(link = "log"),data = mtpl)
AIC_sev_gam_16 <- AIC(sev_gam_16) #AIC = 302805 this model is a better fit

#sev_gam_17 <- gam(amount ~ s(age) + sex + s(codposs) + agecar + fuelc + split + fleetc + coverp + powerc +  agecar:fuelc + age:sex #+ sex:powerc + age:codposs, method = "REML", family =Gamma(link = "log"), data = mtpl)
#AIC_sev_gam_17 <- AIC(sev_gam_17) #  model 16 is a better fit
#AIC_sev_gam_17 

sev_gam_18 <- gam(amount ~ s(age) + sex + s(codposs) + agecar + fuelc + split + fleetc + coverp + powerc +  agecar:fuelc + age:sex + sex:powerc + age:split, method = "REML", family =Gamma(link = "log"),data = mtpl)
AIC_sev_gam_18 <- AIC(sev_gam_18) #AIC = 302798 this model is a better fit

sev_gam_19 <- gam(amount ~ s(age) + sex + s(codposs) + agecar + fuelc + split + fleetc + coverp + powerc +  agecar:fuelc + age:sex + sex:powerc + age:split + age:agecar, method = "REML", family =Gamma(link = "log"),data = mtpl)
AIC_sev_gam_19 <- AIC(sev_gam_19) #AIC = 302783 this model is a better fit

sev_gam_20 <- gam(amount ~ s(age) + sex + s(codposs) + agecar + fuelc + split + fleetc + coverp + powerc +  agecar:fuelc + age:sex + sex:powerc + age:split + age:agecar + fuelc:coverp, method = "REML", family =Gamma(link = "log"),data = mtpl)
AIC_sev_gam_20 <- AIC(sev_gam_20) #AIC = 302786 model 19 is a better fit

x <- c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,18,19,20)
AIC_sev_gam <- data.frame(model_name = numeric(0), model = numeric(0), AIC = numeric(0))
for(i in x){
  AIC_sev_gam[i,] <- c(paste("sev_gam_",i, sep =""), 0 ,round(AIC(get(paste("sev_gam_",i, sep =""))),1) ) 
}
```
To get pure premium we need to combine frequency and severity. We need to calculate expected frequency and expected severity and then multiply them. We will use model with the lowest AIC as it is the best fit. For frequency this is `freq_gam_19` and for severity `sev_gam_19`. Now lets calculate their expected values with function `predict`

`freq_gam_19 <- gam(nclaims ~ s(age) + sex + s(codposs) + agecar + fuelc + split + fleetc + coverp + powerc +  agecar:fuelc + age:sex + sex:powerc + age:split + age:agecar, method = "REML", family =poisson(link = "log"),data = mtpl)`
`sev_gam_19 <- gam(amount ~ s(age) + sex + s(codposs) + agecar + fuelc + split + fleetc + coverp + powerc +  agecar:fuelc + age:sex + sex:powerc + age:split + age:agecar, method = "REML", family =Gamma(link = "log"),data = mtpl)`

```{r, echo=F}
#expected_frequency <- mean(predict(freq_gam_19, newdata = NULL, 
#       type = "response"))
#expected_severity <- mean(predict(sev_gam_19, newdata = NULL, 
#       type = "response"))
#c("Expected frequency" = expected_frequency, "Expected severity" = expected_severity)
```
We see that expected frequency is equal to 0.1239788 and expected sevirty 2527.4901645 EUR (or some other currency). If we multiply this two values we will get the pure premium that we are looking for
```{r, echo=F}
#pure_premium = expected_frequency * expected_severity
#c("Pure premium" = pure_premium)
```




To compare the models
```{r, echo=F}

```

