---
title: "Non-Life Insurance - Final project"
output: html_document
author: "Klemen Hovnik and Manca Strgar"
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages,  include=FALSE, warning = FALSE, message = FALSE}
#package instalation:
packages <- c("tidyverse", "mgcv", "evtree", "classInt", "rgdal", "RColorBrewer", "grid", "gridExtra", "visreg", "sf", "tmap", "rgeos", "mapview", "leaflet", "rmarkdown", "ggplot2")
suppressMessages(packages <- lapply(packages, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x)
    library(x, character.only = TRUE)
  }
  else {
  message("Everything is set up correctly. You are ready to go.")
}
}))

#libraries:
library(dplyr)
library(knitr)
library(rmarkdown)
library(tidyverse)
library(ggplot2)
library(graphics)
library(here)
library(gridExtra)
library(mgcv)
library(evtree)
library(classInt)
library(rgdal)
library(RColorBrewer)
library(grid)
library(visreg)
library(sf)
library(tmap)
library(rgeos)
library(mapview)
library(leaflet)

```
PART 1
=======

# 1 Importing dataset

First, we need to import data from the file `Assignment.csv` into R. The file contains 163.657 rows with 16 variables. The column `nbrtotc` shows total number of claims during the period of exposure. 
Let's have a look at our dataset:

```{r, echo=F}
mtpl_orig <- read.csv("Assignment.csv", header = TRUE)
mtpl_orig=as_tibble(mtpl_orig)
kable(head(mtpl_orig))
#str(mtpl_orig)
#summary(mtpl_orig)

```

We could rename colums in our dataset to make it easier to work with. We will rename the column `nbrtotc` into `nclaims` as the number of claims, and `duree` into `expo` as exposure. For easier programming we will also rename `ageph` and `sexp` into `age` and `sex`. We will also delete colums which are not important for our analysis rightnow. 
```{r, echo=F}
mtpl <- mtpl_orig %>%
  # rename all columns 
  rename_all(function(.name) {
    .name %>% 
      # replace all names with the lowercase versions
      tolower 
    })
mtpl <- rename(mtpl, age= ageph, sex = sexp, nclaims = nbrtotc, expo = duree,
               amount = chargtot)
kable(head(mtpl))
```

# 2 Empirical analysis
For the begining of our empirical analysis, we will calculate the empirical claim frequency. Empirical claim frequency rate is the anticipated percentage of insured that will make claims against insurance and the number of claims they will make during a certain period. SO we calculate it as 
$$emperical\ claim\ frequency = \frac{\sum{number\ of\ claims}}{\sum{exposure}}$$

```{r, echo=F}
mean(mtpl$nclaims)
sum(mtpl$nclaims)/sum(mtpl$expo)
mtpl %>% summarize(emp_freq = sum(nclaims) / sum(expo)) 
```
We can see that aour empirical claim frequency is low, so we will not expect a large number of claims.

To have a better understanding of our data we can make a graphic presentation of our data.
One column in our data represents the number of claims for every individual in our portfolio. This is a very important information for building our model. We can observe that in majority of the cases we have zero claims per policy. We can also compute the mean and variance that will confirm that.

```{r, echo = F, warning=FALSE}
mean1 <- mean(mtpl$nclaims)
var1 <-  sum((mtpl$nclaims - mean1)^2) /length(mtpl$nclaims)
c(mean = mean1, variance = var1)

```
We see that mean is really close to zero. But to convince you even more, that the majority of the policies have zero claims, we can show you in a graph. Graph below actually shows us that almost 150,000 of policies have zero claims and the number of policies that have more than zero claims is verry low.
```{r, echo=F}
graph1<- ggplot(mtpl, aes(nclaims)) + geom_bar(col = "red", 
              fill = "red", alpha = 0.5) + 
     labs(y = "Abs frequency (in exposure)") +
     ggtitle("MTPL - number of claims")
graph1
```

Looking at this in relative terms, we can see that more than around 90% of insurance contract did not have any claims. So approximatelly 10 % of all policyholders have one or more claims.

```{r, echo=F}
graph2 <- ggplot(mtpl, aes(nclaims)) + theme_bw()
graph2 + geom_bar(aes(y = (..count..)/sum(..count..)), 
    col = "red", fill = "red", alpha = 0.5) + 
  labs(y = "Relative frequency") +
  ggtitle("MTPL - relative number of claims")
```

But the number of claims is not the only information that we have in our data. We can also graphically show relative frequency of exposure (`expo`). We can calculate empirical claim frequency, per unit of exposure and variance. TLE NE VEM LIH TOČNO KAJ ZDEJ TO POMEN

```{r first-risk-calculations-mtpl-2, echo = F, warning=FALSE}
mean2 <- sum(mtpl$nclaims) / sum(mtpl$expo); 
var2 <- sum((mtpl$nclaims-mean2*mtpl$expo)^2)/sum(mtpl$expo) 
#mean2
c(mean = mean2,   variance = var2)
#mtpl %>% summarize(emp_freq = mean2, variance = var2 )
```

If we do the same for each gender we see that claim frequency is higher for males than females. ???A NI LIH OBRATNO???

```{r, echo = F, warning=FALSE, message=FALSE}
mtpl %>%  group_by(sexp) %>% summarize(emp_freq = sum(nclaims) / sum(expo))

```
Now we will graphically show side by side relative frequency for our given data.


```{r, echo=F}
ylab <- "Relative frequency"
ggplot.bar <- function(DT, variable, xlab){
  ggplot(data = DT, aes(as.factor(variable)), environment = environment()) + theme_bw() + 
    geom_bar(aes(y = (..count..)/sum(..count..)), col = "red", fill = "red") + labs(x = xlab, y = ylab)
}
ggplot.hist<- function(DT,variable,xlab,binwidth){
  ggplot(data =DT, aes(variable))+theme_bw()+
    geom_histogram (aes(y =(..count..)/sum(..count..)), binwidth = binwidth, col ="red", fill ="red", alpha = 0.5)+
    labs(x=xlab, y=ylab)
} 

sevirity_plot <- ggplot(data = mtpl, aes(mtpl$amount)) + geom_density(adjust = 3, col = "red", fill = "red") + xlim(0,1e4) + ylab(ylab) + xlab("sevirity") + theme_bw()
sevirity_plot
grid.arrange(ggplot.bar(mtpl, mtpl$nclaims, "nclaims"),
ggplot.hist(mtpl, mtpl$expo, "expo", NULL),
sevirity_plot,
ggplot.bar(mtpl, mtpl$coverp, "coverage"),
ggplot.bar(mtpl, mtpl$fuelc, "fuel"),
ggplot.bar(mtpl, mtpl$sex, "sex"),
ggplot.bar(mtpl, mtpl$usec, "use"),
ggplot.bar(mtpl, mtpl$fleetc, "fleet"),
ggplot.hist(mtpl, mtpl$age, "age", NULL),
ggplot.bar(mtpl, mtpl$powerc, "power"),
ggplot.bar(mtpl, mtpl$agecar, "car age"),
ggplot.bar(mtpl, mtpl$sportc, "sport car"), ncol =4)
```
Among our data, we have a information about the postal code of Belgium in which the claim happened. We can show this in a map.

```{r, echo = F, warning=FALSE, message=FALSE}

mtpl_pc <- aggregate(mtpl$expo ~ mtpl$codposs, data=mtpl, sum)
names(mtpl_pc)[names(mtpl_pc) == "mtpl$expo"] <- "EXP"
names(mtpl_pc)[names(mtpl_pc) == "mtpl$codposs"] <- "PC"
mtpl_pc$N <- mtpl_pc$EXP
head(mtpl_pc)

library(rgdal)
setwd('C://Users/manca/Documents/NL')
readShapefile = function(){
  belgium_shape <- readOGR(dsn = path.expand(paste(getwd(), "/Non-Life-insurance", sep = "")), layer = "npc96_region_Project1.shp")
  belgium_shape <- spTransform(belgium_shape, CRS("+proj=longlat +datum=WGS84"))
  belgium_shape$id <- row.names(belgium_shape)
  return(belgium_shape)
}

belgium_shape <- readShapefile()




library(sf)
belgium_shape_sf <- st_read('./Non-Life-insurance/npc96_region_Project1.shp', quiet = TRUE)
belgium_shape_sf <- st_transrofm(belgium_shape_sf, "+proj=longlat +datum=WGS84")
class(belgium_shape_sf)
```





# 3 The construction of a (technical) tariff structure
```{r, echo=F}

mtpl %>% group_by(ageph) %>% 
  summarize(emp_freq = sum(nclaims) / sum(expo)) %>% 
  ggplot(aes(x = ageph, y = emp_freq)) + theme_bw() +
  geom_point(color = "#003366")

glm.freq1 <- glm(nclaims ~ fuelc, offset = log(expo), data = mtpl, family = poisson(link = "log"))
summary(glm.freq1)

gam.freq1 <- gam(nclaims ~ fuelc, offset = log(expo), data = mtpl, family = poisson(link = "log"))
summary(gam.freq1)

gam.freq2 <- gam(nclaims ~ s(ageph), offset = log(expo), data = mtpl, family = poisson(link = "log"))
summary(gam.freq2)

pred <- predict(gam.freq2, type = "terms", se = TRUE)
str(pred)

b <- pred$fit[,1]
l <- pred$fit[,1] - qnorm(0.975) * pred$se.fit[,1]
u <- pred$fit[,1] + qnorm(0.975) * pred$se.fit[,1]
x <- mtpl$ageph
df <- unique(data.frame(x, b, l, u))
df


p <- ggplot(df, aes(x = x))
p <- p + geom_line(aes(y = b), size = 1, col="#003366")
p <- p + geom_line(aes(y = l), size = 0.5, linetype = 2, col="#99CCFF")
p <- p + geom_line(aes(y = u), size = 0.5, linetype = 2, col="#99CCFF")
p <- p + xlab("ageph") + ylab(expression(hat(f)(ageph))) + theme_bw()
p

ggplot.gam <- function(model,variable,gam_term,xlabel,ylabel){
  pred <- predict(model, type = "terms", se = TRUE)
  col_index <- which(colnames(pred$fit)==gam_term)
  x <- variable
  b <- pred$fit[, col_index]
  l <- pred$fit[, col_index] - qnorm(0.975) * pred$se.fit[, col_index]
  u <- pred$fit[, col_index] + qnorm(0.975) * pred$se.fit[, col_index]
  df <- unique(data.frame(x, b, l, u))
  p <- ggplot(df, aes(x = x))
  p <- p + geom_line(aes(y = b), size = 1,col="#003366")
  p <- p + geom_line(aes(y = l), size = 0.5, linetype = 2,col="#99CCFF")
  p <- p + geom_line(aes(y = u), size = 0.5, linetype = 2,col="#99CCFF")
  p <- p + xlab(xlabel) + ylab(ylabel) + theme_bw()
  p
}

plot.gam.freq.ageph <- ggplot.gam(gam.freq2, mtpl$ageph, "s(ageph)", "ageph", expression(hat(f)(ageph)))
plot.gam.freq.ageph


```

