---
title: "Final project - part 2"
output: html_document
author: "Klemen Hovnik and Manca Strgar"
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages,  include=FALSE, warning = FALSE, message = FALSE}
#package instalation:
packages <- c("tidyverse", "here", "gridExtra", "grid", "rstudioapi", "MASS", "actuar", "statmod", "ReIns", "pscl", "rmarkdown", "ggplot2")
new_packages <- packages[!(packages %in% installed.packages()[,"Package"])]
if(length(new_packages)) install.packages(new_packages)

if(sum(!(packages %in% installed.packages()[, "Package"]))) {
  stop(paste('The following required packages are not installed:\n', 
             paste(packages[which(!(packages %in% installed.packages()[, "Package"]))], collapse = ', ')));
} else {
  message("Everything is set up correctly. You are ready to go.")
}
#libraries:
library(dplyr)
library(knitr)
library(MASS)
library(rmarkdown)
library(tidyverse)
library(ggplot2)
library(pscl) 
library(actuar)
library(graphics)
library(survival)
```
# 1 Fitting exponential function

First we need to import our data from the file `SeverityCensoring` into R. The file contains 9062 claims, with columns representing policy id, claim id, rc (iz claim is censored - number or uncensored - value NA ), deductible (always 100) and the claim amount.
```{r, echo = F, warning=FALSE}
Severity <- read.table(file = "./SeverityCensoring.txt", header = T, sep = " ")
```
Before we start fitting exponential distribution to our data we will modify our rc column, so that it is a logical colum (indicator), where TRUE means the claim is censores and FALSE means that the claim is not censored. This is how we do this:
```{r, warning=FALSE}
Severity$rc <- !is.na(Severity$rc) 
```

Now we can finally fit exponential distribution to our model.We will compute negative log-likelihood function, taking trucation and censoring into account. We will do this with the help of `dexp` and `pexp` function that represent density and distribution functions of exponential distribution. As our paramether $\lambda > 0$ we use $\lambda = e^{\beta}$  and now it can take all values.   
```{r, warning=FALSE}
claimAmount <- Severity$claimAmount
rc <- Severity$rc
deductible <- Severity$deductible

exp.negloglikelihood <- function(par)
  {lambda <- exp(par)
  -sum(dexp(claimAmount[!rc],lambda,log=T)) - #here we take values of claimAmount if rc false
    sum(pexp(claimAmount[rc],lambda,log.p=T,lower.tail=F)) + #claimamount if rc true
    sum(pexp(deductible,lambda,log.p=T,lower.tail=F)) #adds deductible
}
```
To get the initial estimation of our parameter, we use the moment estimation. As we know $$ E(X) = \frac1\lambda $$ and from here we get that $$ \lambda = \frac1{m_1}. $$ Where $m_1$ is first moment derived from our data. Now we can fit exponential distribution and we get the MLE estimation of $\lambda$:
```{r, echo = F, warning=FALSE}
par.init <- 1/mean(claimAmount)
fit_exp <- nlm(exp.negloglikelihood,log(par.init))
par.exp <- exp(fit_exp$estimate) 
par.exp
```
The value of AIC is equal to: 
```{r, echo = F, warning=FALSE}
AIC.exp <- 2+2*fit_exp$minimum 
AIC.exp
```


# 2 fitting other distributions
Now for the second pard we need to also fit three other distributions to our data: lognormal, inverse Gaussian and Burr distribution

## 2.1 Fitting lognormal distribution
We repeat our process from above. But we need different inital estimations for our two parameters $\mu$ and $\sigma^2$. We need to use the first and the second moments:
$$ E(X) = e^{ \mu + \frac{\sigma^2}{2}} $$
$$ E(X^2) = e^{2\mu + 2\sigma^2} $$
Now we get first and the second moment from our data and from that our estimations. 
We fit the model and get our estimates for $\mu$ and $\sigma^2$:
```{r, echo = F, warning=FALSE}
lnorm.negloglikelihood <- function(par){
  mu <- exp(par[1]); 
  sig <- exp(par[2])
      -sum(dlnorm(claimAmount[!rc],mu,sig,log=T)) -         
      sum(plnorm(claimAmount[rc],mu,sig,log.p=T,lower.tail=F)) +
      sum(plnorm(deductible,mu,sig,log.p=T,lower.tail=F))
}
m1 <- mean(claimAmount); m2 <- mean(claimAmount^2)
par.init <- c(log(m1^2/sqrt(m2)), log(m2/m1^2))

fit_lnorm <- nlm(lnorm.negloglikelihood,log(par.init))

par.lnorm <- exp(fit_lnorm$estimate) 
par.lnorm
```
And AIC for this fit is:
```{r, echo = F, warning=FALSE}
AIC.lnorm <- 4+2*fit_lnorm$minimum # 121207.2
AIC.lnorm
```

## 2.2 Fitting inverse Gaussian distribution
Here our first two moments, that are used for the initial estimation, are equal to:
$$E(X) = \mu $$ 
$$ E(X^2) = \mu^2 + \frac{\mu^3}{\theta} $$
Whit this we get our estiamtions for $\mu$ and $\theta$:
```{r, echo = F, warning=FALSE}
invG.negloglikelihood <- function(par){
  mu <- exp(par[1]);
  theta <- exp(par[2])
      - sum(dinvgauss(claimAmount[!rc],mu,theta,log=T)) -         
      sum(pinvgauss(claimAmount[rc],mu,theta,log.p=T,lower.tail=F)) +
      sum(pinvgauss(deductible,mu,theta,log.p=T,lower.tail=F))
}
m1 <- mean(claimAmount); m2 <- mean(claimAmount^2)
par.init <- c(m1, m1^3/m2)

fit_invG <- nlm(invG.negloglikelihood,log(par.init))

par.invG <- exp(fit_invG$estimate)
par.invG
```
AIC for inverse Gaussian fit is:
```{r, echo = F, warning=FALSE}
AIC.invG <- 4+2*fit_invG$minimum #120651.7
AIC.invG
```
## 2.3 Fitting Burr distribution

Here we have three paramethers, so we need first three moments, for initial estimation.
$$E(X) = \frac{\theta * \Gamma(1+ \frac{1}{\gamma})* \Gamma(\alpha-\frac{1}{\gamma})}{\Gamma(\alpha)} $$ 
$$E(X^2) = \frac{\theta^2 * \Gamma(1+ \frac{2}{\gamma})* \Gamma(\alpha-\frac{2}{\gamma})}{\Gamma(\alpha)} $$ 
$$E(X^3) = \frac{\theta^3 * \Gamma(1+ \frac{3}{\gamma})* \Gamma(\alpha-\frac{3}{\gamma})}{\Gamma(\alpha)} $$ 
Whit this we get our estiamtions for $\alpha$, $\theta$ and $\gamma$:
```{r, echo = F, warning=FALSE}
neg.loglik.burr <- function(par){
  s1 <- exp(par[1]); s2 <- exp(par[2]); sc <-exp(par[3])
  -sum(dburr(claimAmount[!rc],s1,s2,sc,log=T)) - sum(pburr(claimAmount[rc],s1,s2,sc,log.p=T,lower.tail=F)) +
sum(pburr(deductible,s1,s2,sc,log.p=T,lower.tail=F))
}
par.init <- c(1,1,1)
fit_burr<- nlm(neg.loglik.burr,log(par.init))
par.burr <- exp(fit_burr$estimate)
par.burr 
```
Corresponding AIC is:
```{r, echo = F, warning=FALSE}
AIC.burr <- 6+2*fit_burr$minimum 
AIC.burr
```

# 3 Fitting Erlang mixture distribution
Here we use the felp of EM algorithm for Erlang mixtures which we source from file `EM_MixedErlang`.
We need to define loss and nrc vector so that they meet the conditions form the insutructions. This is how we did this:
```{r, echo = F, warning=FALSE, include=F}
source("EM_MixedErlang.R")
```
```{r, warning=FALSE, message=FALSE}
loss <- claimAmount ; 
nrc <- claimAmount
for (i in 1:9062) {
  if (rc[i] == TRUE){
    nrc[i] <- NA
  }
}
```
Here are the MLE estimations for Erlang mixture distribution:
```{r, echo = F, warning=FALSE}
fit.ME <- ME_fit(loss, nrc, trunclower=100, M=5, s=3)
fit.ME
```

# 4 Plotting the Kaplan-Meier estimate of the survival function
Here we need to define three vectors: deds, loss and full and then plot the fit. We do this like this:
```{r, warning=FALSE}
deds <- deductible ; loss <- claimAmount ; 
full <- rc
for (i in 1:9062) {
  if (rc[i] == TRUE){
    full[i] == FALSE
  }else{full[i] == TRUE}
}
  

fit <- survfit(Surv(deds, loss, full) ~ 1)
plot(fit, mark.time=F, conf.int=F)
```


# 5 Plots
Now we will plot all the survival functions for all the different distributions onto one graph, so that we can compare them. The graph looks like this:

```{r, echo = F, warning=FALSE}

x <- claimAmount

plot(fit, mark.time=F, conf.int=F, add=T)
curve((1-pexp(x,par.exp))/(1-pexp(100, par.exp)), col="red", add=T, ylab ="y") 
curve((1-plnorm(x,par.lnorm))/(1-plnorm(100, par.lnorm)), col="blue", add=T)
curve((1-pinvgauss(x,par.invG))/(1-pinvgauss(100, par.invG)), col="orange", add=T)
curve((1-pburr(x,par.burr[1], par.burr[2], par.burr[3]))/(1-pburr(100, par.burr[1], par.burr[2], par.burr[3])), add=TRUE, col="pink")
curve((1-ME_cdf(x,fit.ME$theta, fit.ME$shape, fit.ME$alpha))/(1-ME_cdf(100,fit.ME$theta, fit.ME$shape, fit.ME$alpha)), add=TRUE, col = "purple")
legend(30000, 0.99, legend=c("Kaplan-Meier", "Exponential", "Lognormal", "Inverse Gaussian", "Burr", "Erlang mixture"), col=c("black", "red", "blue", "orange", "pink", "purple"),lty=1, cex=0.8,
       title="Line types")

```
